Sperm should evolve to make female meiosis fair.

### 2014-05-21 writeup_and_notes/signif_statement.tex
yb [hopefully] improved  gcs sig statement while keeping it at 119 words

\begin{document}
Meiosis is the time when alleles compete to ensure their transmission at the expense of organismal fitness, and many features of meiosis may have evolved to reduce competition. The observation that fertilization is often required for the completion of female meiosis appears incompatible with this view; sperm-acting alleles could distort female meiosis in favor, creating an additional conflict. Counterintuitively, we show that sperm are selected to ensure a fair female meiosis. This comes about because sperm immediately suffer the result of cheating, and so it is in sperm's interests to enforce fairness in female meiosis.  The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
\end{document}

>>>>

\begin{document}
During female meiosis, alleles can take charge of their transmission, and those that bias meiosis in their favor can spread at a cost to organismal fitness. It is therefore believed that basic features of oogenesis likely evolved to control such renegade genes. The observation that female meiosis requires fertilization for its completion appears incompatible with this view; sperm-acting alleles could help like alleles distort  female meiosis in their favor, creating an additional conflict. Counterintuitively, we show that sperm succeed by enforcing fairness in females meiosis because this allows them to avoid pairing with low-fitness cheaters.  The requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
\end{document}
